---
title: "Choosing priors in an ordinal logistic model"
author: "Susannah Tysor"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

    ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE)
```

```{r packages}
library(dplyr)
library(ggplot2)
#library(bayesplot)
library(rstan)
library(cowplot)
library(testthat)
library(parallel)
library(purrr)
```

```{r depends}
source('prior_analysis_functions.R')
```

```{r options}
rstan_options(auto_write=TRUE)
```

## Prior choice

Michael Betancourt thinks that an [induced dirichlet prior](https://betanalpha.github.io/assets/case_studies/ordinal_regression.html) on the cut points may be a good choice for ordinal logistic models.

When I modelled phenology using an ordinal logistic model, I used

* a gamma distribution for the prior on the cut points, $c$
* an exponential distribution distribution for the transition speed/slope, $\beta$
    * normal distributions for group effects on the slope

Normally, priors are bottom up - you choose a prior distribution for each parameter in your model. This is tricky for the cut points in an ordinal logistic model because they're defined on an abstract latent space where your domain expertise doesn't really apply. But you really want to choose a good prior in ordinal logistic models, because ordinal logistic models with covariates are inherently non-identifiable. (Because $\beta$ and $c$ depend on one another.)

Betancourt says that

> To avoid the non-identifiability of the interior cut points from propagating to the posterior distribution we need a principled prior model that can exploit domain expertise to consistently regularize all of the internal cut points at the same time. Regularization of the cut points, however, is subtle given their ordering constraint. Moreover domain expertise is awkward to apply on the abstract latent space where the internal cut points are defined.

He develops an induced dirichlet prior on the cut points. I want to compare the induced dirichlet and gamma priors on cut points for my situation. 

If I simulate data like mine using an ordinal logistic model, can I recapture the parameters using an ordinal logistic model with a gamma or induced dirichlet prior on the cut points?

At this point, I don't actually care about recapturing the cut points or beta. I do care about recapturing the transition points, $h$, though. $h$ are the points at which half of the population has transitioned from one state to another and is calculated as $h = \frac{c}{\beta}$. Eventually I will want to consider how different groups affect $h$ through deviations from the main slope $\beta$ (*e.g* $h = \frac{c}{\beta + \beta_{group_i}}$) and so estimating $\beta$s well might be more important then. But for now, let's just consider a an ordinal logistic model with a covariate/latent effect and no groups.

## Goals
1) When can I recapture parameters with a gamma prior?
2) When can I recapture parameters with an induced dirichlet prior?
3) Which one works best?

# Simulate data
First let's simulate some data that's kind of like mine.

I have three potential states $K$ and a latent effect/covariate $x$. $x$ is always positive and measurements are only made when $x$ is between `8` and `18`. Data collection begins when workers notice trees are approaching state `2` and stops as trees enter state `3`.

I simulate 3 datasets with slow ($\beta = 0.5$), medium ($\beta = 1$), and fast ($\beta = 2$) transition speeds. I want the transitions to occur at the same $x$ for both datasets. The halfway transition points $h$ will be at `x= 10` and `x=15`. $h$ is calculated with cut points $c$ and transition speed $\beta$, that is $h = \frac{c}{\beta}$ So while the transition points will stay the same in each simulated dataset, the cut points will differ. 

Cut points for the slow transition dataset will be at 5 and 7.5, at 10 and 15 for the medium and at 20 and 30 for the fast transition.

Here is what the first transition between states looks like for those transitions. My data is more like the fast transition - a phenological period with transitions of beta=0.5 would be occuring over an unrealistically long time period.

```{r transitions, fig.height=3, fig.width=4}

logistic2 <- function(x, beta, c) {
  y <- 1/(1+exp(-beta*x + c))
}

x <- c(0:20)
logisticexp <- data.frame(x, slow = logistic2(x, 0.5, 5), fast=logistic2(x, 2, 20), medium=logistic2(x, 1,10)) %>%
  tidyr::pivot_longer(cols=c(slow, medium, fast))

ggplot(logisticexp, aes(x=x, y=value, color=name)) +
  geom_line()
```

Now the real fake data:

```{r simulate_data}
simulation_input <- set_simulation_parameters(N=500)

simdat <- map(simulation_input$inputlist, simulate_data, groups=FALSE)

plot_simulated_data(simdat, simulation_input$inputlist)

```

## Attempt to recapture parameters

Now I want to try to recapture parameters.

I need to know how good I have to be at choosing the shape and rate parameters for the gamma prior on the cut points to recapture the "true" slope and cut points parameters as well. (The mean of the gamma is the shape divided by the rate.)  

Choosing where to center the gamma isn't obvious since cut points are on that "abstract latent space." Since $c = \beta h$, though, we have some idea of what that is. $x=12.5$ is smack dab in the middle of our transition points $h$, which we can roughly estimate from the data. It would be nice to be able to use this number that we know when choosing the parameters for the gamma prior. 

If we center a gamma distribution on state `2`, how much wiggle room does it need to find the cut points? When $\beta=0.5$, `12.5` is `6.25` on the cut point space and when $\beta=2$, `12.5` is `25` on the abstract latent space. The larger $\beta$ is, the bigger and further apart the cut points are. The smaller $\beta$ is, the smaller and closer together the cut points are. Since we don't know $\beta$ in advance, we probably have to choose a fat gamma so we can access the full range of values where cut points could be. On the other hand, we might end up estimating combinations of $c$ and $\beta$ that get $h$ right, but both $c$ and $\beta$ very wrong.

I also want to understand how the exponential prior on $\beta$ affects the ability of the model to recapture parameters. I'll try rates of 1,2,3.


### Gamma prior with covariate
```{r paramsForGammaModels}

beta_rate <- c(1:3) # rate parameters for exponential prior on beta

# shape and rate parameters for gamma prior on cut points, all centered on mean transition point

factors <- list(fat=0.10, middling=0.25, skinny=1)

pars_gamma <- purrr::map(factors, make_gamma_pars, h=simulation_input$h)  

#plot gamma priors
pars_gamma %>%
  # format for plotting
  map(function(x) {rgamma(1000, x$shape, x$cut_rate)}) %>%
  bind_cols() %>%
  tidyr::pivot_longer(cols = names(factors)) %>%
  # plot
  ggplot(aes(x=value, fill=name)) +
  geom_density(alpha=0.5) +
 # facet_grid(name ~ .) +
  ggtitle("Gamma priors all centered on 10", subtitle = "with different shape and rate params")

# make a nice dataframe with all combinations params used to simulate data and model params used to try to recover those params
parframe_gam <- pars_gamma %>%
  purrr::map_dfr(.f=bind_cols, .id=".id") %>%
  merge(beta_rate) %>%
  dplyr::rename(gammaid=.id, beta_rate=y) %>%
  merge(simulation_input$pars)
parframe_gam$h1 <- simulation_input$h[1]
parframe_gam$h2 <- simulation_input$h[1]
parframe_gam$modelid <- 1:nrow(parframe_gam) #label the models
```

So there are 3 simulated datasets - 1 for each of three transition speeds. I'm going to try to fit them both with 3 rates for the $\beta$'s exponential prior and 3 shape and rate combinations for the cut points' gamma prior, a total of 27 model runs. 

```{r paramsTableGamma}
knitr::kable(parframe_gam, caption="Model parameterizations used to try to recapture params")
```


```{r GammaRecapture, include=FALSE, cache=TRUE}
# Parallel code actually executes in parallel when running this chuck as a chunck with ctrl+shift+enter. It does not if you run it line by line. I don't know why. 

# format parframe so it works with parLapply better
parlist_gam <- make_parframe_list(parframe_gam)


# run all models, parallelized

# make a cluster using half your cores
no_cores <- parallel::detectCores()/2
cl <- parallel::makeCluster(no_cores)

# export the stuff you need to run on the cluster
parallel::clusterExport(cl, c("fit_gamma_model", "parlist_gam", "simdat"))
parallel::clusterEvalQ(cl, c(library(rstan), library(StanHeaders)))

# fit the models
fits_gam <- parallel::parLapply(cl, parlist_gam, function(x) {fit_gamma_model(simdatlist = simdat, pars=x, groups=FALSE)})

parallel::stopCluster(cl) #close the cluster

```
### Does Stan throw warnings?
Do a very un-thorough check for model problems, like Rhat and divergences.
```{r modelcheck}
map(fits_gam, function(x) {rstan:::throw_sampler_warnings(x)})
```

```{r gammaExtract}

# extract parameters from stan objects pair with true
params_gam <- bind_true_model_pars(fits_gam, parlist_gam)

```

```{r gammaParPlots, fig.height=2}
#plot params and diffs
#map(paramsgam, parplot)
# map(paramsgam, posterior_differencer) %>%
#     map2(.y=paramsgam, .f=diffplotter)
```

### Recapture

Of the 5 parameters ($\beta$, $c$ and derived $h$), how many are recaptured by the model?
```{r GammaHDPI}
# calculate whether true value is in HPDI

in50 <- map(params_gam, calc_HPDI, prob=0.5)

in75 <- map(params_gam, calc_HPDI, prob=0.75)

in90 <- map(params_gam, calc_HPDI, prob=.90)

# recaptured parameters
perform50 <- map_df(in50, bind_rows, .id="modelid") %>%
  dplyr::mutate(modelid=as.integer(modelid)) %>%
  full_join(parframe_gam) 
perform90 <- map_df(in90, bind_rows, .id = "modelid") %>%
  dplyr::mutate(modelid=as.integer(modelid)) %>%
  full_join(parframe_gam)

# proportion of parameters recaptured
prop_recaptured50 <- perform50 %>%
  group_by(modelid) %>%
  summarise(captured = sum(inint)) %>%
  full_join(parframe_gam) %>%
  arrange(beta, desc(captured))

prop_recaptured90 <- perform90 %>%
  group_by(modelid) %>%
  summarise(captured = sum(inint)) %>%
  full_join(parframe_gam) %>%
  arrange(beta, desc(captured))

#knitr::kable(prop_recaptured50, caption = "Proportion of params recaptured in 50% HPDI")
#knitr::kable(prop_recaptured90, caption = "Proportion of params recaptured in 90% HPDI")
```

```{r plotGammaRecaptureoverall}
ggplot(prop_recaptured50, aes(x=gammaid, y=as.factor(beta_rate), fill=captured)) + 
  geom_tile(colour="white") +
  geom_text(aes(label=modelid)) +
  facet_wrap("transition", scales="free") +
  ggtitle("Proportion of parameters recaptured in 50% HPDI", subtitle = "faceted by transition speed in simulated dataset") +
  scale_fill_viridis_b(limits=c(0,5)) +
  ylab("beta_rate")

ggplot(prop_recaptured90, aes(x=gammaid, y=as.factor(beta_rate), fill=captured)) + 
  geom_tile(colour="white") +
  geom_text(aes(label=modelid)) +
  facet_wrap("transition", scales="free") +
  ggtitle("Proportion of parameters recaptured in 90% HPDI",subtitle = "faceted by transition speed in simulated dataset") +
  scale_fill_viridis_b(limits=c(0,5)) +
  ylab("beta_rate")
```

No model recaptures all parameters, but all models return 4/5 of the parameters in the 90% HDPI. Parameters are recaptured better (*i.e.* they show up in the 50% HPDI) when the underlying transition speed is medium ($\beta=1$), the $\beta$ prior is not too loose ($beta_rate=1$ can cause problems), and the gamma shape is fat to middling (don't constrain the gamma prior with a tight distribution).

Which parameters are recaptured?



```{r plotGammaRecapture}
#knitr::kable(perform50, caption = "Ability of the different models to recapture parameters in 50% HPDI. Each model only run only once.")

ggplot(perform50, aes(x=params, y=as.factor(modelid), fill=inint)) + 
  geom_tile(color="black") +
  scale_fill_viridis_d()  +
  geom_text(aes(label=paste(gammaid, beta_rate, sep="\n")), size=2.5) +
  facet_wrap("beta", scales="free") +
  ggtitle("Parameters recaptured in 50% HPDI", subtitle = "faceted by transition speed (beta) in simulated dataset\n with gamma shape and beta rate labels") +
  theme(legend.position = "top")
  

#knitr::kable(perform90, caption = "Ability of the different models to recapture parameters in 90% HPDI. Each model only run only once, don't shoot me.")

ggplot(perform90, aes(x=params, y=as.factor(modelid), fill=inint)) + 
  geom_tile(color="black") +
  scale_fill_viridis_d()  +
  geom_text(aes(label=paste(gammaid, beta_rate, sep="\n")), size=2.5) +
  facet_wrap("beta", scales="free") +
  ggtitle("Parameters recaptured in 90% HPDI", subtitle = "faceted by transition speed (beta) in simulated dataset\n with gamma shape and beta rate labels") +
theme(legend.position = "top")

```

All parameters except the second transition point ($h2$) are recaptured in the 90% HPDI. 

$h$ is *never* recaptured in the 50% HPDI, and, while $h1$ is recaptured in all 90% HPDIs, $h2$ never is. That's a bummer, since that's what I actually want to know. Also, how can $c.2$ and $\beta$ be captured, but $h2$ not be when $h2 = \frac{c.2}{\beta}?