---
title: "Choosing priors in an ordinal logistic model"
subtitle: "with covariate and group effects"
author: "Susannah Tysor"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

```{r}
library(dplyr)
library(rstan)
library(tidyr)
library(purrr)
library(cowplot)
library(bayesplot)

rstan_options(auto_write=TRUE)
```
# Background

[Previously](https://scisus.github.io/ordinal_logistic_prior/prior_analysis.html), I considered whether a gamma or an induced dirichlet prior on cut points in an ordinal logistic model worked better for the kind of data I'm considering.

In this model, the state of flowering depends on how many forcing units have accumulated. 

There are three possible states: 1-not yet flowering, 2- flowering, 3- done flowering. 

$x$ is accumulated forcing: when flowering was observed, how much warmth had the population/tree been exposed to since January 1? $x$ is always positive (and always increases monotonically through time - though time is abstracted out of this model).

$\beta$ describes how fast the transition occurs - small $\beta$s make the transition from between states occur over a wider range of $x$'s. (Translated from forcing units to days, this answers a question like "does the population transition between states over 1 day, 3 days, a week?" We work in forcing unit space because the trees respond to temperature, not time - and no spring heats up exactly the same so dates are kind of useless for prediction.) $\beta$ is always positive.

A gamma prior on cut points that's got a distribution fat enough to encompass the entire possible range of cut points worked well. An induced dirichlet prior on cut points only works well when it is anchored near the first cut point and when cut points are relatively close together. While the model with an induced dirichlet prior can struggle to recapture $\beta$ and cut points $c$, it is pretty good at capturing the relationship between $\beta$ and $c$: $h = \frac{c}{\beta}$, which is the point at which half the trees in a population have transitioned or the point at which an individual tree is 50% likely to have transitioned. 

In the previous analysis, no groups were considered. In reality, I want to know if site, provenance, and clone affect $h$.

## Goals
1 - Generate data with an ordinal logistic model 
 a - With linear model structure $(\beta + \beta_{site_i}) x$
 b - With linear model structure $\beta x + \alpha$
2 - Fit both models
3 - Determine whether $\beta$, \beta_{site_i}$, \alpha$, $c$, or $h$ can be returned.

# Simulate site effects

Let's simulate 7 groups with `N=100` observations for each group. Each group's $h$ are shifted by $h_mod$, which is drawn from a standard normal distribution.

```{r}
hist(rnorm(1000), breaks=20)
```


```{r group simulation}

set_simulation_parameters <- function(N=100, hsd=1) {
    K <- 3
    G <- 7 # number of groups (sites)
    N <- N

    beta <- data.frame(transition = c("medium", "fast"), beta=c(1, 2))
    cutpoints <- data.frame(c.1= c(10, 20), c.2=c(15, 30), transition=c("medium", "fast"))
    
    h_base <- data.frame(h.1 = 10, h.2 = 15) # half transitions
    h_mod <- rnorm(G, 0, hsd) # how much being in each group modifies half transitions
    
    groups <- data.frame(group = 1:G, h_mod = h_mod)

    simu_pars <- merge(beta, cutpoints) %>%
      merge(h_base) %>%
      merge(groups) %>%
      mutate(h.1_group = h.1 + h_mod, h.2_group = h.2 + h_mod) %>%
      mutate(beta_group = (c.1/(h.1_group) - beta), alpha_group = (beta*(h.1_group) + c.1))


    # covariate around active period
    x <- runif(n=N, min=8, max=18)

    inputs_for_sim <- split(simu_pars, simu_pars$transition) %>%
        purrr::map(.f=function(y) {list("N" = N, "K" = K, "G" = G, "c" = unique(c(y$c.1, y$c.2)), "beta"=unique(y$beta), "beta_g" = unique(y$beta_group), "alpha_g" = unique(y$alpha_group), "group" = unique(y$group), "x" = x)})

    return(list(pars=simu_pars, inputlist=inputs_for_sim, h=h_base))
}

simulation_input <- set_simulation_parameters()

# simulate from models with identical transition points, but effects on alpha or beta

simulate_data <- function(input, modification = "beta") {
  # simulate data
  if (effect == "beta") {
    simu <- rstan::stan(file='simulate/covar_group_beta_sim.stan', iter=1, chains=1, algorithm="Fixed_param", data=input)
  } else {
    simu <- rstan::stan(file='simulate/covar_group_alpha_sim.stan', iter=1, chains=1, algorithm="Fixed_param", data=input)
  }
  # extract data from stan model
simu_params <- data.frame(rstan::extract(simu)$y[1,,]) %>%
  mutate(x = input$x) %>%
  pivot_longer(cols = -x, names_to = "group", values_to = "y") %>%
  extract(group, into = "group", regex = "([:digit:])") %>%
  mutate(group = as.integer(group))
  
  # format data as input to another stan model

  input_data_for_model <- list("N" = input$N * input$G, "K" = input$K, "G" = input$G, "x" = simu_params$x, "y" = simu_params$y, "group" = simu_params$group)
  
  return(input_data_for_model)
}


simdat_beta <- map(simulation_input$inputlist, simulate_data, modification = "beta")  
names(simdat_beta) <- paste(names(simdat_beta), "beta")
simdat_alpha <- map(simulation_input$inputlist, simulate_data, modification = "alpha")
names(simdat_alpha) <- paste(names(simdat_alpha), "alpha")
simdat <- list()
simdat <- append(simdat, simdat_beta)
simdat <- append(simdat, simdat_alpha)

# Plot simulated data to make sure it's reasonable
plot_simulated_data <- function(simdat, simulation_input) {
    # format for plotting
    simdf <- purrr::map(simdat, .f = function(x) {x[c("x", "y", "group")]}) %>%
        purrr::map_dfr(.f = bind_rows, .id=".id")

    # scatterplot
    p1 <- ggplot(simdf, aes(x=x, y=y)) +
        geom_jitter(shape=1, height=0.1, alpha=0.5) +
        ggtitle("Simulated data") +
        facet_grid(group ~ .id)

    # cumulative plot
    p2 <- ggplot(simdf, aes(x=x, colour=as.factor(y))) +
        stat_ecdf() +
        theme(legend.position = "none") +
        ggtitle("Cumulative x for 3 states") +
        facet_grid(group ~ .id)

    cowplot::plot_grid(p1, p2, nrow=2)
}

plot_simulated_data(simdat, simulation_input$inputlist)
```

# Recapture

Can the parameter values be recaptured when a model with an induced dirichlet prior for the cut points is used?

```{r params}

beta_rate <- c(1:3) # rate parameters for exponential prior on beta
anchor <- c(5, 10) # different anchor parameters for induced dirichlet prior

# make a nice dataframe with all combinations params used to simulate data and model params used to try to recover those params
# 

transition <- simulation_input$pars %>% 
  select(transition) %>%
  distinct()

parframe_indir <- merge(transition, y=beta_rate) %>%
    rename(beta_rate=y) %>%
    merge(y=anchor) %>%
    rename(anchor=y) 
parframe_indir$modelid <- 1:nrow(parframe_indir)

# format parframe so it works with parLapply better
make_parframe_list <- function(parframe) {
    parlist <- split(parframe, seq(nrow(parframe)))
    names(parlist) <- parframe$modelid

    return(parlist)
}
parlist_indir <- make_parframe_list(parframe_indir)
```

```{r indirRecapture, include=FALSE}
# fit a model with an induced dirichlet prior on cutpoints in stan. simdatlist is a list of simulated data (1 simulated dataset per list entry), pars is a list of parameter values (1 set of parameter values per list entry) and groups is TRUE or FALSE indicating whether you're trying to fit groups.
fit_indir_model <- function(simdatlist, pars, modification) {
    #choose whether to use data simulated with a rapid or slow transition
    if (pars$transition == "medium") {
        simdat <- simdatlist$medium
    }
    if (pars$transition == "fast") {
        simdat <- simdatlist$fast
    }
    #extract parameters for prior distributions
    simdat$anchor <- pars$anchor
    simdat$beta_rate <- pars$beta_rate

    #fit the model
    if (modification == "beta") {
        fitindir <- stan(file='induced_dirichlet/dirichlet_covar_beta_group.stan', data=simdat, chains=4)
    } else {
        fitindir <- stan(file='induced_dirichlet/dirichlet_covar_alpha_group.stan', data=simdat, chains=4)
    }
    return(fitindir)
}

# run all models, parallelized

# make a cluster using half your cores
no_cores <- parallel::detectCores()/2
cl <- parallel::makeCluster(no_cores)

# export the stuff you need to run on the cluster
parallel::clusterExport(cl, c("fit_indir_model", "parlist_indir", "simdat_beta", "simdat_alpha"))
parallel::clusterEvalQ(cl, c(library(rstan), library(StanHeaders)))

reps=10
for (i in 1:reps) {
  fits_indir <- parallel::parLapply(cl, parlist_indir, function(x) {fit_indir_model(simdatlist = simdat_beta, pars=x, modification="beta")})
  saveRDS(fits_indir, file = paste0("induced_dirichlet/group_runs/beta_dat_beta_mod/run", i, ".rds"))
  rm(fits_indir)
  gc()
  
    fits_indir <- parallel::parLapply(cl, parlist_indir, function(x) {fit_indir_model(simdatlist = simdat_beta, pars=x, modification="alpha")})
  saveRDS(fits_indir, file = paste0("induced_dirichlet/group_runs/beta_dat_alpha_mod/run", i, ".rds"))
  rm(fits_indir)
  gc()
  
    fits_indir <- parallel::parLapply(cl, parlist_indir, function(x) {fit_indir_model(simdatlist = simdat_alpha, pars=x, modification="beta")})
  saveRDS(fits_indir, file = paste0("induced_dirichlet/group_runs/alpha_dat_beta_mod/run", i, ".rds"))
  rm(fits_indir)
  gc()
  
    fits_indir <- parallel::parLapply(cl, parlist_indir, function(x) {fit_indir_model(simdatlist = simdat_alpha, pars=x, modification="alpha")})
  saveRDS(fits_indir, file = paste0("induced_dirichlet/group_runs/alpha_dat_alpha_mod/run", i, ".rds"))
  rm(fits_indir)
  gc()
}

parallel::stopCluster(cl) #close the cluster

```

```{r readIndirModels, warning=FALSE, cache=TRUE}

# pull in parameters and info on divergences, etc from saved stanfit objects
# very slow step. consider parallelizing to the extent your ram can handle
# 
## append a label (string) to all columnnames in a dataframe (x)
label_names <- function(x, label) {
    colnames(x) <- paste0(colnames(x), "_", label)
    return(x)
}

#bind true parameters (in list parlist) and model parameters (in list fit) even tho it will make a giant df
bind_true_model_pars <- function(fits, parlist) {
    # extract params from model object
    params <- lapply(fits, function(x) {data.frame(rstan::extract(x) ) } )

    # label params as coming from the model or as true params used to
    params <- map(params, label_names, label="model")
    parlist <- map(parlist, label_names, label="true")

    # combine model and true params in a big list of dataframes - each list entry is a dataframe for a single model
    params <- map2(params, parlist, cbind)

    return(params)
}

# take a stan object and find out if there's anything egregiously wrong with it
check_model <- function(stanobj) {
    nuts <- nuts_params(stanobj)
    divergences <- dplyr::filter(nuts, Parameter=="divergent__" & Value==1) %>%
        nrow()
    rhats <- rhat(stanobj)
    bad_rhats <- sum(rhats > 1.01)
    nefrats <- neff_ratio(stanobj)
    bad_neff <- sum(nefrats < 0.1, is.nan(nefrats), na.rm=TRUE)
    diagnostics <- data.frame(divergences = divergences, bad_rhats=bad_rhats, bad_neff=bad_neff)
    return(diagnostics)
}

# run check_model on a list of models and add a column that names each row by the list name
check_list_of_models <- function(model_list) {
    map_dfr(model_list, check_model, .id=".id") %>%
        rename(modelid=.id)
}

# Extract model configurations and the most obvious problems with fit/convergence from stanfit objects. Stanfit objects are in lists in .rds files saved in path. path is a string denoting the directory where the rds files are stored. Nothing other than .rds files with lists of stanfit objects should be stored in path.
extract_pars_and_problems <- function(path, parlist) {
    fits <- list.files(path=path)
    reps <- length(fits)
    pars <- list()
    bad_models <- list()

    for (i in 1:reps) {
        run <- readRDS(paste0(path, fits[i])) # read in first run of 27 models
        run_pars <- bind_true_model_pars(run, parlist=parlist) # extract parameters
        run_fit <- check_list_of_models(run) %>% # id problems
            mutate(all_bads = divergences + bad_rhats + bad_neff) %>%
            filter(all_bads > 0) %>%
            select(-all_bads)
        pars[[i]] <- run_pars
        bad_models[[i]] <- run_fit
        rm(run, run_pars, run_fit)
        gc()
    }

    return(list(pars = pars, problems = bad_models))
}

extracts_indir_beta_alpha <- extract_pars_and_problems(path="induced_dirichlet/group_runs/beta_dat_alpha_mod/", parlist=parlist_indir)
extracts_indir_beta_beta <- extract_pars_and_problems(path="induced_dirichlet/group_runs/beta_dat_beta_mod/", parlist=parlist_indir)
```


```{r indirModelCheck}
problems_indir_ba <- extracts_indir_beta_alpha$problems %>%
  map_dfr(bind_rows, .id=".id") %>%
  rename(run=.id) %>%
  group_by(modelid) %>%
  summarize(bad_proportion = n()/reps, bad_count = n(), divergences_mean=mean(divergences), bad_rhats_mean=mean(bad_rhats), bad_neff_mean=mean(bad_neff))
  
knitr::kable(problems_indir_ba, caption = "Models with divergences, rhats > 1, or neff_ratios < 0.1 - induced dirichlet")

problems_indir_bb <- extracts_indir_beta_beta$problems %>%
  map_dfr(bind_rows, .id=".id") %>%
  rename(run=.id) %>%
  group_by(modelid) %>%
  summarize(bad_proportion = n()/reps, bad_count = n(), divergences_mean=mean(divergences), bad_rhats_mean=mean(bad_rhats), bad_neff_mean=mean(bad_neff))

knitr::kable(problems_indir_bb, caption = "Models with divergences, rhats > 1, or neff_ratios < 0.1 - induced dirichlet")

```
